|

### 2021-08-18 appendix/appendix.Rmd
improving section 1 appendix

The city council votes the mayor's proposal and if they accept it, it is implemented. Otherwise, a reversal policy takes place. Reversal policy in this context means that the mayor's proposal failed to secure majority of council votes. The decision leaves the mayor's hands, and either immediately stops or it is transferred to the council. The types of reversal policies are crucial for our argument, as it changes the relative strength of the councilors in their bargaining with the mayor. We investigate three types of reversal policies: a \textit{non-partisan} reversal mechanism, where city councilors have no partisan concerns; a \textit{partisan} reversal mechanism, where the council is divided in supporters and non-supporters of the mayor; and finally a \textit{hybrid} reversal mechanism, that combines partisan concerns with non-partisan rent extraction. In the main paper, we focus on the hybrid model, as it is the more realistic scenario. Despite that, comparing the partisan and non-partisan reversal mechanisms illustrate clearly our argument that the partisan a mechanism is what drives our results.

To govern, the mayor has to convince at least half of the city councilors to support her policy proposals. Similar to selectorate models \citep{buenodemesquita2005logic}, we assume that the selectorate, i.e. the individuals that have a say in the mayor's policy proposals, are the $N$ city councilors. The winning coalition, which is the minimum number of members that have to be convinced to implement the policy, is equal to half of the selectorate $N/2$. City councilors are moved by policy concerns $p$ and transfers received from the mayor \textbf{x}. The use of transfers to affect the councilors' choices generate a cost of govern, $C_G$, that vary with the reversal mechanism and the legislature size.

Finally, after paying the costs of govern, the mayor proposes a level of public goods provision ($g$) and rents ($r$). Public goods provision help the mayor to get reelected. Rents are for direct mayor's consumption, and does not contribute to the mayor's electoral success. This makes the mayor's expected utility a sum of the gains from rents and the benefits from reelection. Both the utilities from rents $u(r)$ and the probability of reelection $\pi(g)$ are concave functions, meaning that more rents or public goods increase the utility at a decreasing rate.\footnote{In technical terms, $u'>0$ and $\pi'>0$; and $u''<0$ and $\pi''<0$.} The probability of reelection is multiplied by the benefit from holding office $B_M > 0$. This benefit captures the tangible and intangible gains that the mayor perceives from holding the public office. We assume that the benefits are high enough to rule out an equilibrium with only rents and transfers. If the mayor's policy proposal is approved, then the expected utility for the mayor is:

@@

The expected utility for the city councilors depend on the reversal policy. As discussed, we explore three different mechanisms. First, in \textit{non-partisan} reversal mechanism, when the council rejects the mayor's proposal, a reversal stage starts with resources diminish by a factor of $\delta \in (0,1)$ and the selection of one councilor randomly. The randomly selected councilor becomes now the proposer. If her proposal is accepted, it is implemented. If it is rejected, the budget shrinks again by a factor $\delta$, and another city councilor is recognized to make the proposal. The process repeats, until one proposal is finally accepted.

In the second reversal mode, the \textit{partisan} reversal mechanism, we assume that each councilor has a party affiliation. Party affiliations are mutually exclusive (a councilor cannot belong to two parties at once). If a city councilor is affiliated to a party aligned with the mayor, we say that she belongs to the government coalition $G \subset \{1, 2, \cdots, N \}$. Otherwise, the city councilor belongs to the opposition $O \not\subset G$. The mayor's policy generate a political value of $p > 0$ for the councilors, and whether it adds or subtracts from the councilor's utility depends of the party affiliation of the councilor. If the mayor's proposal is rejected, a city councilor aligned with the mayor, that would receive $p$ from the policy, now receives zero. On the other hand, a city councilor that belongs to the opposition would receive $-p$ from the mayor's policy, and now receives zero. Although these assumptions are strong, they capture the underlying political dynamics of partisanship: the political preferences in terms of public goods provision are shared by councilors aligned with the mayor. Voting against the mayor's proposals would represent to defy the partisan brand.

>>>>

The city council votes the mayor's proposal and if they accept it, it is implemented. Otherwise, a reversal policy takes place. Reversal policy in this context means that the mayor's proposal failed to secure majority of council votes. The decision process leaves the mayor's office, and either immediately stops or it is transferred to the council. The types of reversal policies are crucial for our argument, as it changes the relative strength of the council in their bargaining with the mayor. We investigate three types of reversal policies: a \textit{non-partisan} reversal mechanism, where city councilors have no partisan affiliation; a \textit{partisan} reversal mechanism, where the council is divided in supporters and non-supporters of the mayor; and finally a \textit{hybrid} reversal mechanism, that combines partisan with non-partisan features. In the main paper, we focus on the hybrid model, as it is the more realistic scenario. Despite that, comparing the partisan and non-partisan reversal mechanisms illustrate clearly our argument that the partisan a mechanism is what drives our results.

To govern, the mayor has to convince at least half of the city councilors to support her policy proposals. Similar to \citet{buenodemesquita2005logic}, we assume that the selectorate, i.e. the individuals that have a say in the mayor's policy proposals, are the $N$ city councilors. The winning coalition, which is the minimum number of members that have to be convinced to implement the policy, is equal to half of the selectorate $N/2$. City councilors are moved by policy concerns $p$ and transfers received from the mayor or the councilor recognized as the proposer \textbf{x}. We leave the definition of these transfers open, as this facilitates the relation with several instruments that mayors can use to pressure city councilors. To illustrate, examples of transfers are granting electoral favors, spending in areas that the councilor has political or personal interests, granting portfolio and public jobs within the municipal bureaucracy, or even campaign contributions and bribes. The use of transfers to affect the councilors' choices and generate a governing cost of $C_G$. The governing costs vary with the reversal mechanism and the legislature size.

Finally, after paying the costs of govern, the mayor proposes a level of public goods provision ($g$) and rents ($r$). Public goods provision help the mayor to get reelected. Rents are for direct mayor's consumption, and does not contribute to the mayor's electoral success. This makes the mayor's expected utility a sum of the gains from rents and the benefits from reelection. Both the utilities from rents $u(.)$ and the probability of reelection $\pi(.)$ are concave functions, meaning that more rents or public goods increase the utility at a decreasing rate.\footnote{In technical terms, $u'>0$ and $\pi'>0$; and $u''<0$ and $\pi''<0$.} The probability of reelection is multiplied by the benefit from holding office $B_M > 0$. This benefit captures the tangible and intangible gains that the mayor perceives from holding the public office. We assume that the benefits are high enough to rule out an equilibrium with only rents and transfers, which is equivalent to say that mayors have a strong preference to be reelected. If the mayor's policy proposal is approved, then the expected utility for the mayor is:

@@

The expected utility for the city councilors depend on the reversal policy. As discussed, we explore three different mechanisms. First, in \textit{non-partisan} reversal mechanism, when the council rejects the mayor's proposal, a reversal stage starts with resources diminish by a factor of $\delta \in (0,1)$ and the selection of one councilor randomly. The randomly selected councilor becomes the new proposer. If her proposal is accepted, it is implemented. If it is rejected, the budget shrinks again by a factor $\delta$, and another city councilor is recognized to make the proposal. The process repeats, until one proposal is finally accepted.

In the second reversal mode, the \textit{partisan} reversal mechanism, we assume that each councilor has a party affiliation. Party affiliations are mutually exclusive (a councilor cannot belong to two parties at once). If a city councilor is affiliated to a party aligned with the mayor, we say that she belongs to the government coalition $G \subset \{1, 2, \cdots, N \}$. Otherwise, the city councilor belongs to the opposition $O  = \{1, 2, \cdots, N \} \setminus G$. The mayor's policy generate a political value of $p > 0$ for the councilors, and whether it adds or subtracts from the councilor's utility depends of the party affiliation of the councilor. If the mayor's proposal is rejected, a city councilor aligned with the mayor, that would receive $p$ from the policy, now receives zero. On the other hand, a city councilor that belongs to the opposition would receive $-p$ from the mayor's policy, and now receives zero. Although these assumptions are strong, they capture the underlying political dynamics of partisanship: the political preferences in terms of public goods provision are shared by councilors aligned with the mayor. Voting against the mayor's proposals would represent to defy the partisan label.
